---
title: "0-dataImport"
author: "Will Sheppard"
date: "2024-04-18"
output: html_document
---

THROUGHOUT THIS SCRIPT I USE "OF" AS SHORT HAND FOR OPTIC FLOW AS THIS WAS THE NAME OF THE TASK WHEN I WROTE THE SCRIPTS INITIALLY

YOU CAN RUN THE SCRIPT AS A WHOLE AND THE INDIVIDUAL DATA FRAMES WILL BE GENERATED, BUT YOU WILL NEED TO GO THROUGH AND RUN THE SAVE CHUNKS MANUALLY. THIS IS TO STOP ACCIDENTAL OVER SAVING

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
This function takes a list of the required packages, checks if they are installed, downloads those that are not, and then loads them.

```{r packages, warning=FALSE}

requiredPackages = c("tidyverse", "rstatix")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

import some custom funcitons that i have written to work with the vision tests
```{r funcitons}

# import functions
source('functions_visionProcessing_and_threshEstimation.R')
```

task contains oex3

```{r import task data}

# create list of file names
files <- c("task-d4fw", "task-dqjp", "task-f3ud", "task-mva3", "task-oex3",
           "task-og71", "task-rek9", "task-s8ch", "task-svji", "task-z5dj")
for(i in files){
  filepath <- file.path("../data/",paste("data_exp_62188-v20_",i,".csv",sep=""))
  assign(i, read.csv(filepath))
}

# make reaction time variables numeric
`task-d4fw`$Reaction.Time <- as.numeric(`task-d4fw`$Reaction.Time)
`task-d4fw`$X.Coordinate <- as.numeric(`task-d4fw`$X.Coordinate)
`task-d4fw`$Y.Coordinate <- as.numeric(`task-d4fw`$Y.Coordinate)
`task-d4fw`$Timed.Out <- as.numeric(`task-d4fw`$Timed.Out)

taskData <- bind_rows(`task-d4fw`, `task-dqjp`, `task-f3ud`, `task-mva3`, `task-oex3`,
                     `task-og71`, `task-rek9`,`task-svji`, `task-z5dj`)

# remove gaps from column names
names(taskData) <- make.names(names(taskData), unique = TRUE)

# remove participant 5384850 - not sure which condition is which
# leaves n of 59
taskData <- taskData %>%
  filter(Participant.Private.ID != 5384850)

# create a condition column 
taskData <- taskData %>%
  mutate(condition = case_when(
    Response == "Monocular blur" ~ "blur",
    Response == "No blur" ~ "clear"
  ))
# fill column with condition above
taskData <- taskData %>% fill(condition)

allTasks <- taskData
```

let's save this out for future use
```{r save all data, eval=FALSE}

write.csv(allTasks, "../savedData/allTask.csv")

```


Now, let's seperate the heading data and calculate click angle
```{r seperate heading data}

# select data for ofTest and ofCalibrate
ofTasks <- c("ofTest", "ofCalibrate")
responseZones <- c("response_button_text", "response_text_entry", "click_painting")

ofAllData <- allTasks %>%
  filter(display %in% ofTasks) %>%
  filter(Zone.Type %in% responseZones)%>%
  filter(Response != "click limit reached")

## Calculate angle to the centre for all data points
# create center height variable
centerHeight <- 255

ofAllData <- ofAllData %>%
  mutate(centerDist = X.Coordinate - 320)%>%
  mutate(centerAng = atan(centerDist/centerHeight)*(180/pi))
```

Seperate the calibration data
```{r seperate calibration data}
## select ofCalibrate data
allOfCalibrate <- ofAllData %>%
  filter(display == "ofCalibrate")

# select columns
ofCalibrate <- allOfCalibrate %>%
  select(Participant.Private.ID, condition, randomise_trials, cross, X.Coordinate, centerDist, centerAng)

# read in calibration target location spreadsheets
calibTargets <- read_csv("../calibrationTargets.csv")

# create new column with angle to center
calibTargets <- calibTargets %>%
  mutate(targetCenterAng = atan(centerDist/Y.Coordinate)*(180/pi))


# add new column matching target angle to cross
ofCalibrate <- ofCalibrate %>%
  mutate(targetCenterAng = case_when(
    cross == calibTargets$cross[1] ~ calibTargets$targetCenterAng[1],
    cross == calibTargets$cross[2] ~ calibTargets$targetCenterAng[2],
    cross == calibTargets$cross[3] ~ calibTargets$targetCenterAng[3],
    cross == calibTargets$cross[4] ~ calibTargets$targetCenterAng[4],
    cross == calibTargets$cross[5] ~ calibTargets$targetCenterAng[5]
  )) %>%
  mutate(angDiff = abs(targetCenterAng - centerAng))

```

check if any particiapnts need excluding based upon their calibration score
- this happens for both the still and moving calibration tasks
- refer to moving calibrations as "practice"

These participants performance in the main task will be checking script 3. if their performance is sig worse than the mean, they will be removed.
```{r check calibation exclusion}

# By particiapnt and condition, look at mean error
ofCalibrateSummary <- ofCalibrate %>%
  group_by(Participant.Private.ID, condition)%>%
  summarise(meanError = mean(angDiff))

calibEx <- mean(ofCalibrateSummary$meanError)+2*sd(ofCalibrateSummary$meanError)

ofCalibrateExclude <- ofCalibrateSummary %>%
  filter(meanError > calibEx)%>%
  group_by(Participant.Private.ID)%>%
  summarise(count = n())%>%
  filter(count == 2)

###### Now analyse second practice data
# create list of practice videos
practiceVids <- c("40.mp4", "40right.mp4")

allOfPractice <- ofAllData %>%
  filter(video %in% practiceVids)

allOfPractice <- allOfPractice %>%
  mutate(targetCenterAng = case_when(
    video == "40.mp4" ~ -40,
    video == "40right.mp4" ~ 40
  )) %>%
  mutate(angDiff = abs(targetCenterAng - centerAng))

# create summary measures
ofPracticeSummary <- allOfPractice %>%
  group_by(Participant.Private.ID, condition)%>%
  summarise(meanError = mean(angDiff))

pracEx <- mean(ofPracticeSummary$meanError)+2*sd(ofPracticeSummary$meanError)

ofPracticeExclude <- ofPracticeSummary %>%
  filter(meanError > pracEx)%>%
  group_by(Participant.Private.ID)%>%
  summarise(count = n())%>%
  filter(count == 2)

# create lists of participants to exclude
calibExcludeList <- ofCalibrateExclude$Participant.Private.ID
pracExcludeList <- ofPracticeExclude$Participant.Private.ID
excludeList <- unique(c(calibExcludeList, pracExcludeList))

# how many people did we exclude?
print(paste("We may exclude ", length(excludeList), " participants", sep = ""))

# filter data removing excluded participats
# ofFiltData <- ofAllData %>%
#   filter(!Participant.Private.ID %in% excludeList)
```


here, we create a list of videos that were used to in the main heading perception task

We will also create a list of the remaining pps in the heading task. this can be used to help filter participants later on
```{r seperate heading perception data}

# create list of videos to exclude
exVids <- c("40.mp4", "40right.mp4", "")

ofMain <- ofAllData %>%
  filter(!video %in% exVids)

# create target, signed error, absolute error, offset and contrast variables
allOfMain <- ofMain %>%
  mutate(contrast = case_when(
    str_detect(video, "^50*") ~ 50,
    str_detect(video, "^90*") ~ 10,
    TRUE ~ 100
  )) %>%
  mutate(offset = case_when(
    str_detect(video, "4.mp4$") ~ -4, str_detect(video, "4right.mp4$") ~ 4,
    str_detect(video, "8.mp4$") ~ -8, str_detect(video, "8right.mp4$") ~ 8,
    str_detect(video, "12.mp4$") ~ -12, str_detect(video, "12right.mp4$") ~ 12,
    str_detect(video, "16.mp4$") ~ -16, str_detect(video, "16right.mp4$") ~ 16,
    str_detect(video, "20.mp4$") ~ -20, str_detect(video, "20right.mp4$") ~ 20,
    TRUE ~ 0
  )) %>%
  mutate(angError = centerAng-offset)%>%
  mutate(absAngError = abs(angError))

# create a  list of remaining participants in optic flow
ofPps <- unique(ofMain$Participant.Private.ID)

```

let's save this out for future use
```{r save heading perception task data, eval=FALSE}

write.csv(allOfMain, "../savedData/allOfMain.csv")
write.csv(ofPPs, "../savedData/allOfMain.csv")

```

extract the cs and va data from the general task data frame
```{r seperate vision Data}

# create variable with Zone.Name s of inerest
zones <- c("response_button_text", "response_text_entry", "click_painting")

# filter data into individual data sets for cs, va and of
csAllData <- allTasks %>%
  filter(Zone.Type %in% zones)%>%
  filter(display == "cs") 

vaAllData <- allTasks %>%
  filter(Zone.Type %in% zones)%>%
  filter(display == "va")
```

create summary data frames for modelling
"blur"  "clear"
```{r process cs data}

## Process cs Data

# select columns of interest
csData <- csAllData %>%
  select(Participant.Private.ID, condition, Screen.Name, Attempt, Correct, Reaction.Time)%>%
  rename(id = Participant.Private.ID)


# 5372572- spamming the same key on blur CS - remove
csData <- csData %>%
  filter(id != 5372572)
csPps <- unique(csData$id)

csBlurData <- csData %>%
  filter(condition == "blur")
csClearData <- csData %>%
  filter(condition == "clear")

csBlurShort <- shortenCS(csBlurData)
csBlurSumm <- summShortCS(csBlurShort)
csBlurSumm$condition <- "blur"

csClearShort <- shortenCS(csClearData)
csClearSumm <- summShortCS(csClearShort)
csClearSumm$condition <- "clear"

csSumm <- bind_rows(csBlurSumm, csClearSumm)
```

```{r save csSumm, eval=FALSE}

write.csv(csSumm, "../savedData/csSumm.csv")
```

```{r process va data}

## Process va Data

# select columns of interest
vaData <- vaAllData %>%
  select(Participant.Private.ID, condition,Screen.Name, Attempt, Correct, Reaction.Time)%>%
  rename(id = Participant.Private.ID)

vaPps <- unique(vaData$id)


vaBlurData <- vaData %>%
  filter(condition == "blur")
vaClearData <- vaData %>%
  filter(condition == "clear")

vaBlurShort <- shortenVA(vaBlurData)
vaBlurSumm <- summShortVA(vaBlurShort)
vaBlurSumm$condition <- "blur"

vaClearShort <- shortenVA(vaClearData)
vaClearSumm <- summShortVA(vaClearShort)
vaClearSumm$condition <- "clear"

vaSumm <- bind_rows(vaBlurSumm, vaClearSumm)
```

```{r save vaSumm, eval=FALSE}

write.csv(vaSumm, "../savedData/vaSumm.csv")
```

```{r check all participants are the same}

visionPps <- intersect(csPps, vaPps)
pps <- intersect(ofPps, visionPps)

ofOut <- allOfMain %>%
  filter(Participant.Private.ID %in% pps)
csOut <- csSumm %>%
  filter(id %in% pps)
vaOut <- vaSumm %>%
  filter(id %in% pps)

```


```{r save all data, eval=FALSE}

write.csv(excludeList, "../savedData/ofPoorCalibPPs.csv")
write.csv(ofOut, "../savedData/allOfMain.csv")
write.csv(pps, "../savedData/pps.csv")
write.csv(csOut, "../savedData/csSumm.csv")
write.csv(vaOut, "../savedData/vaSumm.csv")
```