---
title: "1-csProcessing"
author: "Will Sheppard"
date: "2024-04-18"
output: html_document
---
5 particiapnts removed due to cs task

1 participants is removed for repeatedly pressing "s" in response (in last script)
4 participants removed for giving data that cannot be analysed

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries
This function takes a list of the required packages, checks if they are installed, downloads those that are not, and then loads them.

```{r packages, warning=FALSE}

requiredPackages = c("tidyverse", "quickpsy")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

read data 
5384675 - removed due to responding letters not on in the task even at very high contrast
```{r read data}
data <- read.csv("../savedData/csSumm.csv") %>%
  filter(id != 5384675)

```


The Bland-Altman analysis had the greatest levels of agreement when CS thresholds were estimated with a Weibull function and a 1/26 lapse rate. 
```{r fit psychometric function}



conditionList <- c("blur", "clear")
funList <- "weibull_fun"
denomList <- 26

cols <- c("id",
          "cs", "csLower", "csUpper",
          "threshold", "threshLower", "threshUpper",
          "percCorrect","fun", "guessRate",
          "condition")



for (cond in conditionList){

  conditionData <- data %>%
    filter(condition == cond)
      
      pplist <- unique(conditionData$id)
      #pplist <- 1:10
      
      thresh <- data.frame(matrix(
      data = NA,
      nrow = length(pplist),
      ncol = 11))
      
      colnames(thresh) <- cols

      
  for (fun in funList) {
    for (denom in denomList) {
      for (pp in  pplist){
      
          d <- conditionData %>%
            filter(id == pp)
        
          
          
          g = 1/denom
          p <- g + .5 * (1 - g)
          
          xMAX <- max(d$contrast)
          xMIN <- min(d$contrast)
          minPercCorr = min(d$percCorrect)
          minContrast = min(d$contrast)
          
          print(paste("Participant:", pp, " ", cond, "", ", ", fun, "", ", Guess rate: 1/", denom, sep = ""))
          
          if (minPercCorr < p) {
            if (fun == "logistic_fun"){
              fitOne <- quickpsy(d = d, x = contrast, k = totalCorrect, n = totalAttempt,
                             guess = g, lapses = TRUE, bootstrap = 'nonparametric',
                             xmin = xMIN, xmax = xMAX, fun = logistic_fun)
            }else if (fun == "cum_normal_fun"){
              fitOne <- quickpsy(d = d, x = contrast, k = totalCorrect, n = totalAttempt,
                             guess = g, lapses = TRUE, bootstrap = 'nonparametric',
                             xmin = xMIN, xmax = xMAX, fun = cum_normal_fun)
            }else if (fun == "weibull_fun"){
              fitOne <- quickpsy(d = d, x = contrast, k = totalCorrect, n = totalAttempt,
                             guess = g, lapses = TRUE, bootstrap = 'nonparametric',
                             xmin = xMIN, xmax = xMAX, fun = weibull_fun)
            }
            
            x <- plotcurves(fitOne)+
            scale_x_continuous(limits = c(0, max(d$contrast)), breaks = seq(0, max(d$contrast), 1))+
            scale_y_continuous(limits = c(0,1), breaks = seq(0, 1, 0.1))+
            ggtitle(paste("pp: ", pp, " - ", fun, "guess rate 1/", denom))
            show(x)
          
            thresh[which(pplist == pp), 1] <- pp
            thresh[which(pplist == pp), 2] <- round(2 + log10(1/fitOne$thresholds[1]), digits = 2)
            thresh[which(pplist == pp), 3] <- round(2 + log10(1/fitOne$thresholds[4]), digits = 2)
            thresh[which(pplist == pp), 4] <- round(2 + log10(1/fitOne$thresholds[3]), digits = 2)
            thresh[which(pplist == pp), 5] <- round(fitOne$thresholds[1], digits = 2)
            thresh[which(pplist == pp), 6] <- round(fitOne$thresholds[3], digits = 2)
            thresh[which(pplist == pp), 7] <- round(fitOne$thresholds[4], digits = 2)
            thresh[which(pplist == pp), 8] <- fitOne$thresholds[2]
            thresh[which(pplist == pp), 9] <- fun
            thresh[which(pplist == pp), 10] <- 1/denom
            thresh[which(pplist == pp), 11] <- cond
            
          } else{
              thresh[which(pplist == pp), 1] <- pp
              thresh[which(pplist == pp), 2] <- round(2 + log10(1/minContrast), digits = 2)
              thresh[which(pplist == pp), 3] <- round(2 + log10(1/minContrast), digits = 2)
              thresh[which(pplist == pp), 4] <- round(2 + log10(1/minContrast), digits = 2)
              thresh[which(pplist == pp), 5] <- minContrast
              thresh[which(pplist == pp), 6] <- minContrast
              thresh[which(pplist == pp), 7] <- minContrast
              thresh[which(pplist == pp), 8] <- p
              thresh[which(pplist == pp), 9] <- fun
              thresh[which(pplist == pp), 10] <- 1/denom
              thresh[which(pplist == pp), 11] <- cond
          }
          
        
      }
      assign(paste(fun, denom, "cs", cond, sep = ""), thresh)
      }
  
}
  
}

csThresh <- bind_rows(weibull_fun26csblur, weibull_fun26csclear)
#colnames(csThresh) <- cols
```
	
5496748 - clear
5623534 - blur
5549866 - blur
```{r remove inf pps}

ppsExclude <- c(5496748, 5623534, 5549866)

csThresh <- csThresh %>%
  filter(!id %in% ppsExclude)

```



save data
```{r save cs data}

write.csv(csThresh, "../savedData/csThreshold.csv")
```