---
title: "3-ofProcessing"
author: "Will Sheppard"
date: "2024-04-18"
output: html_document
---

4 participants with visual conditions were removed as they performed sig worse than the other participants
Of the 3 participants identified as poor in practce, 2 have already been removed and the remaining 1 had sig higher error so was removed.

TOTAL OF 5 PPS REMOVED

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read in data 
libraries
This function takes a list of the required packages, checks if they are installed, downloads those that are not, and then loads them.

```{r packages, warning=FALSE}

requiredPackages = c("tidyverse", "quickpsy")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

read data 
```{r read data}
csData <- read.csv("../savedData/csThreshold.csv")
vaData <- read.csv("../savedData/vaThreshold.csv")
ofDataAll <- read.csv("../savedData/allOfMain.csv")
#poorCalibPps <- as.list(read.csv("../savedData/ofPoorCalibPPs.csv"))
poorCalibPps <- c(5722070, 5372572, 5384675)
```

we lost 2 participants in cs data processing. double check that all participants line up across the tasks
```{r check all participants are the same}

csPps <- unique(csData$id)
vaPps <- unique(vaData$id)

visionPps <- intersect(csPps, vaPps)


ofPpsData <- ofDataAll %>%
  filter(Participant.Private.ID %in% visionPps)

```

as there are 4 particiapnts with visual conditions that may need to be removed, we will start with a 'gentle' approach to cleaning where we will only remove outliers for specific conditions - group by id, conditions, contrast and offset
```{r remove outliers}


# moving through 4 layers of filtering to get rid of outliers. 
# first remove general outliers
# then get rid of outliers on the individual level (no task component)
# get rid of outliers on a task level (no indivisual component)
# get rid of any remaining outliers on a micro level
ofOutData <- ofPpsData %>% 
  # mutate(outlier = is_outlier(absAngError))%>% # create logical markers for general outliers. this will get rid of the very extreme points
  # mutate(outlier = as.numeric(outlier))%>%
  # filter(outlier == 0)%>%
  # group_by(Participant.Private.ID)%>% # this should catch any very strange behaviours at an individual level
  # mutate(outlier = is_outlier(absAngError))%>% # 
  # mutate(outlier = as.numeric(outlier))%>%
  # filter(outlier == 0)%>%
  # group_by(condition, contrast, offset)%>% # this will catch strange behaviours at a task level
  # mutate(outlier = is_outlier(absAngError))%>%
  # mutate(outlier = as.numeric(outlier))%>%
  # filter(outlier == 0) %>%
  group_by(Participant.Private.ID, condition, contrast, offset)%>% # this will catch strange behaviours al the microlevel
  mutate(outlier = is_outlier(absAngError))%>%
  mutate(outlier = as.numeric(outlier))%>%
  filter(outlier == 0) %>%
  ungroup()

nrow(ofOutData)/nrow(ofPpsData)
```

```{r check visual conditions}

## Check participants with medical condition vs general trend
# 5508881 - cataracts
# 5532853 - Deuteranopia (color blind)
# 5540429 - had cataract, still has film over left eye
# 5658787 - Diabetic Retinopathy
visProbs <- c(5508881, 5532853, 5540429, 5658787)

# create data frame where participants with visual issues are marked
probData <- ofOutData %>%
  mutate(prob = case_when(
    Participant.Private.ID %in% visProbs ~ 1,
    TRUE ~ 0
  ))
# look at descriptive stats for the groups
library(rstatix)

# obviously higher
probData %>%
  group_by(prob) %>%
  get_summary_stats(absAngError, type = "mean_sd")

# calculate t score
prob.test <- probData %>%
  t_test(absAngError ~ prob) %>%
  add_significance()
prob.test

# participants with self reported visual problems had a sig. higher absAngError
# therefore remove
ofOutData <- subset(ofOutData, !(Participant.Private.ID %in% visProbs))

# create a  list of remaining participants in optic flow
#ofPps <- unique(ofOutData$Participant.Private.ID)
```


check the performance of poor calibration pps compared to the remaining pps 
poorCalibPps
```{r poor calib pps}

# create data frame where participants with visual issues are marked
calibData <- ofOutData %>%
  mutate(calib = case_when(
    Participant.Private.ID %in% poorCalibPps ~ 1,
    TRUE ~ 0
  ))

# obviosuly worse
calibData %>%
  group_by(calib) %>%
  get_summary_stats(absAngError, type = "mean_sd")

# calculate t score
prob.test <- calibData %>%
  t_test(absAngError ~ calib) %>%
  add_significance()
prob.test

exPps <- calibData %>% 
  filter(calib == 1)
length(unique(exPps$Participant.Private.ID))

ofOutData <- subset(ofOutData, !(Participant.Private.ID %in% poorCalibPps))


```

save data
```{r save all data, eval=FALSE}


write.csv(ofOutData, "../savedData/ofClean.csv")

```
