---
title: "5-csMLM"
author: "Will Sheppard"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read in data 
libraries
This function takes a list of the required packages, checks if they are installed, downloads those that are not, and then loads them.

```{r packages, warning=FALSE}

requiredPackages = c("tidyverse", "quickpsy", "rstatix")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

read data 
```{r read data}

ofData <- read.csv("../savedData/ofClean.csv")
ofPps <- unique(ofData$Participant.Private.ID)
csAll <- read.csv("../savedData/csThreshold.csv")
cs <- csAll %>%
  filter(id %in% ofPps)
demo <- read.csv("../savedData/demoData.csv")
```

lets check the distributions of CS data
data is approaching normal and doesn't fit alternative distributions
model with all 3 and cg=ompare fit
```{r cs distribution}
ggplot(cs, aes(x = cs, fill = condition))+
  geom_density(alpha = 0.4)

cs %>%
  group_by(condition) %>%
  shapiro_test(cs)
```

To identify the reserarcher, we need to identify the 3 unique combos of OS and monitor size
```{r researcher id}

resDF <- demo %>%
  dplyr::select(Participant.Private.ID, Participant.OS, Participant.Monitor.Size) 

resDF <- na.omit(resDF)
  
resDF <- resDF %>%
  mutate(pcID = paste(Participant.OS, Participant.Monitor.Size)) %>%
  mutate(resID = case_when(
    pcID == "Windows 10 1536x864" ~ 1,
    pcID == "Windows 10 1280x720" ~ 2,
    pcID == "Mac OS 10.15.7 1440x900" ~ 3
  )) %>%
  dplyr::select(Participant.Private.ID, resID)%>%
  rename(id = Participant.Private.ID)

```

Create a data frame containing participant id and their age
```{r particiapnt age}

age <- demo %>%
  select(Participant.Private.ID, age_in_years) %>%
  rename(id = Participant.Private.ID)

```

now we will join the three data frames on Participant id 
revalue condition so that clear is 0 and blur is 1. the coefficient will down us the 'effect' of blur on va
```{r join df}

csAge <- left_join(cs, age)
d <- left_join(csAge, resDF, by = "id") %>%
  rename(age = age_in_years)%>%
  mutate(condition = case_when(
    condition == 'clear' ~ 0,
    TRUE ~ 1
  ))

d$id <- as.factor(d$id)
d$condition <- as.factor(d$condition)
d$resID <- as.factor(d$resID)
d$age.s <- as.numeric(scale(d$age))

```

```{r packages}

requiredPackages = c("lme4", "lmerTest")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

lets start with the gaussian distribution


cs ~ condition*age + (1|resID) + (condition|id) - too many random effects for for number of observations

MAXIMAL MODEL: cs ~ condition*age + resID + (1|resID) + (1|id) 

MINIMAL MODEL: cs ~ condition*age + (1|resID) + (1|id)
```{r cs gausian mlm, eval=FALSE}

# gauMax <- lmer(cs ~ condition*age + #FIXED
#                  (1|resID) + (condition|id), #RANDOM
#          data = d)

# bobyqa - failed to converge
# Nelder_Mead - failed ot converge
# nloptwrap - failed ot converge
# gauMax <- lmer(cs ~ condition*age.s + resID + #FIXED
#                  (1|resID) + (1|id), #RANDOM
#          data = d, 
# #         control = lmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 10000000))
# )
# summary(gauMax)

# bobyqa - failed to converge
# Nelder_Mead - failed ot converge
# nloptwrap - failed ot converge
# gauTwo <- lmer(cs ~ condition*age.s + resID + #FIXED
#                  (1|id), #RANDOM
#          data = d,
#          control = lmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 10000000))
#          )
# summary(gauTwo)

# bobyqa - converged. all results significant
# Nelder_Mead - 
# nloptwrap - 
gauThree <- lmer(cs ~ condition + age.s + resID + #FIXED
                 (1|id), #RANDOM
         data = d, 
         control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000))
         )
summary(gauThree)

gauFinal <- gauThree
```

```{r inverse gausian}

# control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
# bobyqa - failed to converge
# Nelder_Mead - failed to converge
# nloptwrap - failed ot converge
# invGauMax <- glmer(cs ~ condition*age.s + resID + #FIXED
#                   (1|resID) + (1|id), #RANDOM
#                 family = inverse.gaussian(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 10000000))
#          )
# summary(invGauMax)

# bobyqa - failed to converge
# Nelder_Mead - converged. no sig interaction
# nloptwrap - failed ot converge
# invGauTwo <- glmer(cs ~ condition*age.s + resID +#FIXED
#                   (1|id), #RANDOM
#                 family = inverse.gaussian(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
#          )
# summary(invGauTwo)

# bobyqa - failed to converge
# Nelder_Mead - converged. no sig interaction
# nloptwrap - failed ot converge
invGauThree <- glmer(cs ~ condition + age.s + resID +#FIXED
                  (1|id), #RANDOM
                family = inverse.gaussian(link = "identity"),
                data = d,
                control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
         )
summary(invGauThree)

invGauFinal <- invGauThree
```

```{r gamma}

# bobyqa - singular
# Nelder_Mead - singular
# nloptwrap - singular
# gammaMax <- glmer(cs ~ condition*age.s + resID + #FIXED
#                   (1|resID) + (1|id), #RANDOM
#                 family = Gamma(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
#          )
# summary(gammaMax)

# bobyqa - 
# Nelder_Mead - converged, no sig interaction
# nloptwrap - 
# gammaTwo <- glmer(cs ~ condition*age.s + resID + #FIXED
#                   (1|id), #RANDOM
#                 family = Gamma(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
#          )
# summary(gammaTwo)

# bobyqa - 
# Nelder_Mead - converged, no sig interaction
# nloptwrap - 
gammaThree <- glmer(cs ~ condition + age.s + resID + #FIXED
                  (1|id), #RANDOM
                family = Gamma(link = "identity"),
                data = d,
                control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
         )
summary(gammaThree)

gammaFinal <- gammaThree
```


```{r packages}

requiredPackages = c("modelsummary", "tibble", "flextable")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

gamma as lowest BIC
```{r model summary}

modelsummary(list("Inverse Gaussian" = invGauFinal, "Gamma" = gammaFinal), 
             stars = T, 
             title = "CS MLM output",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             fmt = fmt_decimal(digits = 2, pdigits = 3),
             shape = term ~ model + statistic, #statistics in separate columns
             output = "csDistTable.html"
             
             )
```