---
title: "7-ofMLM"
author: "Will Sheppard"
date: "2024-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, warning=FALSE}

requiredPackages = c("tidyverse", "quickpsy", "rstatix")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

read data 
```{r read data}

ofData <- read.csv("../savedData/ofClean.csv")
ofPps <- unique(ofData$Participant.Private.ID)
vaAll <- read.csv("../savedData/vaThreshold.csv")
va <- vaAll %>%
  filter(id %in% ofPps)
csAll <- read.csv("../savedData/csThreshold.csv")
cs <- csAll %>%
  filter(id %in% ofPps)
demo <- read.csv("../savedData/demoData.csv")
```

quite a long tail. may be inv g or gamma
not normally distributed
model using inv g or gamma
```{r of distribution}
ggplot(ofData, aes(x = absAngError, fill = condition))+
  geom_density(alpha = 0.4)

ofData %>%
  group_by(Participant.Private.ID, condition, contrast, offset) %>%
  summarise(error = mean(absAngError))%>%
  ungroup()%>%
  group_by(condition, contrast, offset) %>%
  shapiro_test(error)
```

To identify the reserarcher, we need to identify the 3 unique combos of OS and monitor size
```{r researcher id}

resDF <- demo %>%
  dplyr::select(Participant.Private.ID, Participant.OS, Participant.Monitor.Size) 

resDF <- na.omit(resDF)
  
resDF <- resDF %>%
  mutate(pcID = paste(Participant.OS, Participant.Monitor.Size)) %>%
  mutate(resID = case_when(
    pcID == "Windows 10 1536x864" ~ 1,
    pcID == "Windows 10 1280x720" ~ 2,
    pcID == "Mac OS 10.15.7 1440x900" ~ 3
  )) %>%
  dplyr::select(Participant.Private.ID, resID)%>%
  rename(id = Participant.Private.ID)

```

Create a data frame containing participant id and their age
```{r particiapnt age}

age <- demo %>%
  select(Participant.Private.ID, age_in_years) %>%
  rename(id = Participant.Private.ID)

```

now we will join the three data frames on Participant id 
revalue condition so that clear is 0 and blur is 1. the coefficient will down us the 'effect' of blur on va
```{r join df}

va <- va %>%
  select(id, va, condition)

cs <- cs %>%
  select(id, cs, condition)

vision <- left_join(cs, va)

visionAge <- left_join(vision, age)

extraDf <- left_join(visionAge, resDF, by = "id") %>%
  rename(age = age_in_years)


of <- ofData %>%
  rename(id = Participant.Private.ID) %>%
  select(id, condition, contrast, absAngError)

d <- left_join(of, extraDf, by = c("id", "condition"))%>% 
  mutate(contrast = case_when(
    contrast == 100 ~ "FF",
    contrast == 50 ~ "SDF",
    contrast == 10 ~ "HDF"
  ))

d$id <- as.factor(d$id)
d$condition <- as.factor(d$condition)
d$resID <- as.factor(d$resID)
d$contrast <- as.factor(d$contrast)
d$age.s <- as.numeric(scale(d$age)+3)
d$cs.s <- as.numeric(scale(d$cs)+3)
d$va.s <- as.numeric(scale(d$va)+3)
d$absAngError.t <- d$absAngError+1
d$absAngError.s <- as.numeric(scale(d$absAngError)+3)


levels(d$condition)

clear_v_blur <- c(1/2, -1/2)

contrasts(d$condition) <- cbind(clear_v_blur)

levels(d$contrast)

high_vs_degraded <- c(-2/3, 1/3, 1/3)
low_vs_medium <- c(0, 1/2, -1/2)

contrasts(d$contrast) <- cbind(high_vs_degraded, low_vs_medium)

```

```{r packages}

requiredPackages = c("lme4", "lmerTest")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

MAXIMAL MODEL: absAngError ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
                  (condition|id) + (1|resID) #RANDOM
                  
                  
MINIMAL MODEL: absAngError ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
                  (1|id) #RANDOM

inverse gausian
```{r invGau mlm}

# bobyqa - SINGULAR. resID intercept has the lowest variance associated with it. remove
# Nelder_Mead - 
# nloptwrap - 
# invGauMax <- glmer(absAngError ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
#                      (condition|id) + (1|resID), #RANDOM
#                 family = inverse.gaussian(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000))
#          )
# summary(invGauMax)

# bobyqa - Failed ot converge
# Nelder_Mead - failed to converge
# nloptwrap - 
# invGauTwo <- glmer(absAngError ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
#                      (condition|id), #RANDOM
#                 family = inverse.gaussian(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 10000000))
#          )
# summary(invGauTwo)

# bobyqa - 
# Nelder_Mead - 
# nloptwrap - 
invGauThree <- glmer(absAngError.t ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
                     (1|id), #RANDOM
                family = inverse.gaussian(link = "identity"),
                data = d,
                control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000))
         )
summary(invGauThree)

invGauFinal <- invGauThree
```

gamma
```{r gamma MLM}

# bobyqa - SINGULAR. resID intercept has the lowest variance associated with it. remove
# Nelder_Mead - 
# nloptwrap - 
# gammaMax <- glmer(absAngError.t ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
#                      (condition|id) + (1|resID), #RANDOM
#                 family = Gamma(link = "identity"),
#                 data = d,
#                 control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000))
#          )
# summary(gammaMax)

# bobyqa - converged
# Nelder_Mead - 
# nloptwrap - 
gammaTwo <- glmer(absAngError.t ~ age.s + cs.s + va.s + resID + condition*contrast + #FIXED
                     (condition|id), #RANDOM
                family = Gamma(link = "identity"),
                data = d,
                control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000))
         )
summary(gammaTwo)

gammaFinal <- gammaTwo
```

```{r packages}

requiredPackages = c("modelsummary", "tibble", "flextable")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}

```

gamma as lowest BIC
```{r model summary}

modelsummary(list("Inverse Gaussian" = invGauFinal, "Gamma" = gammaFinal), 
             stars = T, 
             title = "HP MLM output",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             fmt = fmt_decimal(digits = 2, pdigits = 3),
             shape = term ~ model + statistic, #statistics in separate columns
             output = "hpDistTable.html"
             )
```

save data out to make graphs
```{r save data}

write.csv(d, "../savedData/mlm.csv")
```